Friends Get More Money, Attention and Handshakes: Chinese Foreign Aid, Xinhua and Diplomatic Visits

### 2022-08-05 04_paper/Draft4.Rmd
some minor edits to intro

China's overseas development financing now surpasses that of the United States, the World Bank, and several other multilateral development and lending institutions combined. But despite the mamoth proportions of Beijing's aid (and increasingly debt) that it issues world wide, the motives behind its overseas activities remain a mystery. China neither reports its foreign aid spending to international organizations, as do Development Assistance Committee (DAC) members to the OECD, nor does China publish its aid expenditures in a public database, as does the United States in USAID's *Greenbook*. @fuchsRudyak2019 cite this fact as a reason Western governments assume Beijing deliberately keeps its international development finance a secret.

@@

We propose that despite the lack of formal reporting, Beijing can pursue recognition and diplomacy with its overseas financing with the help of other complementary foreign policy tools. Past research has shown that two specific instruments may play special roles in complementing these foreign policy objectives. First, international media are one way that countries can generate attention and promote a positive spin on their activities in foreign audiences. As Nye (2008) notes, daily communications that reach both domestic and foreign populations are a way to improve a country's soft power. Perhaps in recognition of this, Beijing has expanded its outward facing state sponsored media activities and news reporting over the past two decades. This has been done in part through the English version of *Xinhua*, China's top state sponsored news agency.

@@

As we test these hypotheses, in our research design we take care to ensure that we delineate China's development financing according to type. Until relatively recently, many have mistakenly conflated Beijing's ODA-like giving with its other official financing (OOF). However, efforts to disentangle China's activities with respect to grants and low-interest loans (financing that would meet the OECD's official definition of official development assistance or ODA) from its activities with respect to higher interest loans and export credits (OOF) has unmasked a considerable divergence in goals. The conventional view is that while *aid* (ODA) is closely tied to Beijing's foreign policy goals, *debt* (OOF) is related to Beijing's commercial interests.^[See @dreherEtAl2022.] Thus, as we test our hypotheses we also take care to distinguish between Beijing as donor from Beijing as lender.

>>>>

China's overseas development financing now surpasses that of the United States, the World Bank, and several other multilateral development and lending institutions combined. But despite the mammoth proportions of Beijing's aid (and increasingly debt) that it issues world wide, the motives behind its overseas activities remain a mystery. China neither reports its foreign aid spending to international organizations, as do Development Assistance Committee (DAC) members to the OECD, nor does China publish its aid expenditures in a public database, as does the United States in USAID's *Greenbook*. For this reason, many Western governments assume Beijing deliberately keeps its international development finance a secret[@fuchsRudyak2019].

@@

We propose that despite the lack of formal reporting, Beijing can pursue recognition and diplomacy with its overseas financing with the help of other complementary foreign policy activities. Past research has shown that two specific instruments may play special roles in complementing these foreign policy objectives. First, international media are one way that countries can generate attention and promote a positive spin on their activities in foreign audiences. As Nye (2008) notes, daily communications that reach both domestic and foreign populations are a way to improve a country's soft power. Perhaps in recognition of this, Beijing has expanded its outward facing state sponsored media activities and news reporting over the past two decades. This has been done in part through the English version of *Xinhua*, China's top state sponsored news agency.

@@

As we test these hypotheses, in our research design we take care to ensure that we delineate China's development financing according to type. Until relatively recently, many have mistakenly conflated Beijing's ODA-like giving with its other official financing (OOF). However, efforts to disentangle China's activities with respect to grants and low-interest loans (financing that would meet the OECD definition of official development assistance or ODA) from its activities with respect to higher interest loans and export credits (OOF) has unmasked a considerable divergence in goals. The conventional view is that while *aid* (ODA) is closely tied to Beijing's foreign policy goals, *debt* (OOF) is related to Beijing's commercial interests.^[See @dreherEtAl2022.] Thus, as we test our hypotheses we also distinguish between Beijing-as-donor from Beijing-as-lender.
